Homeostatic mechanisms may shape the type and duration of oscillatory modulation.

### 2020-03-06 main.tex

\begin{document}
\maketitle
\linenumbers
% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\begin{abstract}
Neural oscillations are observed ubiquitously in the mammalian brain, but their stability is quite variable. Some oscillations are \textit{tonic} and last for seconds or even minutes. Others oscillations are unstable. They appear for only as a single-cycle \textit{burst}. Likewise, some oscillations rely on excitatory AMPAergic synapses, but others are GABAergic and inhibitory. Why both kinds of diversity are present is not known. We hypothesized Ca$^{2+}$-dependent homeostasis is important to finding an explanation. We tested this in a highly simplified model of hippocampal neurons. In this model homeostasis can profoundly alter the modulatory effect of neural oscillations. Under homeostasis, tonic AMPAergic oscillations actually \textit{decrease} excitability and desynchronize firing. Tonic oscillations that are synaptically GABAergic--like those in real hippocampus--don't provoke a homeostatic response, however. If our simple model is correct, homeostasis can explain why the theta rhythm in hippocampus is synaptically inhibitory: GABA has little to no intrinsic homeostatic response, and so can preserve a cell's dynamic range under tonic oscillation. Based on these results we also speculate that homeostasis can explain why AMPAergic oscillations in cortex often appear as bursts. Bursts minimally interact with the slow homeostasis time constant, and so retain their excitatory effect.\\
\end{abstract}

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\noindent
\textsc{\textbf{New and Noteworthy}} The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known. The interplay between oscillatory modulation and homeostasis is not. We studied oscillatory modulation and intrinsic homeostasis for the first time in a simlified model of hippocampal neurons. We report a paradoxical result: Ca\textsuperscript{2+}-mediated homeostasis causes normally synchronizing AMPAergic oscillations to become inhibitory and desynchronizing. This result, along with other new observations, suggests homeostasis might be just as complex -- and just as important -- for oscillations as it is for all other kinds of modulation.

>>>>

\begin{document}
\maketitle
\linenumbers
% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\begin{abstract}
Neural oscillations are observed ubiquitously in the mammalian brain, but their stability is quite variable. Some oscillations are \textit{tonic} and last for seconds or even minutes. Other oscillations appear as unstable \textit{bursts}. Likewise, some oscillations rely on excitatory AMPAergic synapses, but others are GABAergic and inhibitory. Why all this diversity is present isn't well understood. We hypothesized Ca$^{2+}$-dependent homeostasis is important to finding an explanation. We tested this in a highly simplified model of hippocampal neurons. In this model homeostasis can profoundly alter the modulatory effect of neural oscillations. Under homeostasis, tonic AMPAergic oscillations actually \textit{decrease} excitability and desynchronize firing. Tonic oscillations that are synaptically GABAergic--like those in real hippocampus--don't provoke a homeostatic response, however. If our simple model is correct, homeostasis can explain why the theta rhythm in hippocampus is synaptically inhibitory: GABA has little to no intrinsic homeostatic response, and so can preserve a cell's dynamic range under tonic oscillation. Based on these results we also speculate that homeostasis can explain why AMPAergic oscillations in cortex often appear as bursts. Bursts minimally interact with the slow homeostasis time constant, and so retain their excitatory effect.\\
\end{abstract}

% -----------------------------------------------------
% -----------------------------------------------------
% -----------------------------------------------------
\noindent
\textsc{\textbf{New and Noteworthy}} The intricate interplay of neuromodulators, like acetylcholine, with homeostasis is well known. The interplay between oscillatory modulation and homeostasis isn't. We studied oscillatory modulation and homeostasis for the first time using a simplified model of hippocampus. We report a paradoxical result: Ca-mediated homeostasis causes AMPAergic oscillations to become effectively inhibitory. This result, along with other new observations, suggests homeostasis might be just as complex and important for oscillations as it for other neuromodulators.
